\documentclass[   % 
  final,          % 
  a4paper         % 
]{article}

\usepackage[a4paper,margin=2cm]{geometry}

\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{lightgray}{gray}{0.95}

\usepackage{tikz}
\usetikzlibrary{matrix,calc,shapes}
\usepackage{amssymb}
\usepackage[image]{tikzinput}

\usepackage{standalone}

% Typesetting codeprettier
\usepackage{listings}
\input{lstdefs.tex}

% Break tables
\usepackage{longtable}

% a little better spacing
\usepackage{setspace}
\onehalfspacing
% Symbols for SE and GH
\usepackage{fontawesome} 
% Symbols for creative commons
\usepackage{ccicons}
% linking to the outside world
\usepackage{hyperref}
% Martin's weird SE handle
% (use a different package and Xe/LuaLaTeX if you're doing proper Japanese texts.)
\usepackage{CJKutf8} 

% What this is about
\title{Manual for \texttt{tools-for-g16.bash} (0.3.2, 2020-01-16)}
\author{Martin C Schwarzer}
\date{\today}

\begin{document}
\maketitle

\tableofcontents
\clearpage

\subsection*{Preamble}

The following document is a more verbose version of the already existing reference sheet.
As documentation often goes, it takes some time to reflect all changes properly.
Some sections (especially section~\ref{sec:usage-examples}) may still be incomplete,
a work in progress.
I apologise for the inconvenience.

\input{intro.tex}
\input{install.tex}
\input{utilities.tex}
\input{examples.tex}

\section{License (GNU General Public License v3.0)}

tools-for-g16.bash - A collection of tools for the help with Gaussian 16.
Copyright (C) 2019-2020 Martin C Schwarzer

This program is free software: you can redistribute it and/or modify
it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or
(at your option) any later version.

This program is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
GNU General Public License for more details.

You should have received a copy of the GNU General Public License
along with this program.  If not, see \href{https://www.gnu.org/licenses/}{https://www.gnu.org/licenses/}.

\section{Author, Bugs, and the Rest}

This repository was created and is maintained by Martin C Schwarzer 
(\faStackExchange~\href{https://chemistry.stackexchange.com/users/4945}{Martin-\begin{CJK*}{UTF8}{min}マーチン\end{CJK*}}
  \faGithub~\href{https://github.com/polyluxus}{polyluxus}).

If you find any bugs, have questions, want features implemented,
please do it via the \href{https://github.com/polyluxus/tools-for-g16.bash/issues}{GitHub Issue tracker},
or fork it and send a pull request.

The help of my colleagues for testing, finding bugs, extending the documentation is greatly appreciated.

\begin{flushright}
  This software is licensed GPL-3.0.\\
  This document is licensed \ccbysa.
\end{flushright}

\end{document}



